New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2022-12-29 Manuscript/Manuscript.Rmd
Changes from hackMD

---
title: New insights into the Weddell Sea ecosystem applying a quantitative network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web, which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on the species and trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted food web properties, and analysed species’ roles with respect to the stability of the food web. We found that the distribution of the interaction strength at the food web level is asymmetric, with many weak interactions and few strong ones. We detected a positive relationship between species mean interaction strength and two unweighted properties (i.e., trophic level and the total number of interactions). We also found that only a few species possess key positions in terms of food web stability. These species are charcterised by high mean interaction strength, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In this study, we integrated unweighted and weighted food web information, enabling a more complete assessment of the ecosystem structure and function of the Weddell Sea food web. Our results provide new insights, which are important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

bibliography: ../WeddellSea.bib
running:
  title: New insights into the Weddell Sea food web
  author: Marina et al.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
# copyrightstatement: |
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
codedata: |
    R code and data set used in this work is available at https://github.com/EcoComplex/WeddellSea
#   sample: |
# videosupplement: |
authorcontribution: |
  TIM and LAS: Conceptualization (lead); Data curation (lead); Formal analysis (lead); Methodology (lead); Coding (lead); Writing – original draft (lead); Writing – review and editing (lead). SK: Conceptualization (lead); Formal analysis (supporting); Methodology (supporting); Coding (supporting); Writing – original draft (supporting); Writing – review and editing (supporting).
# disclaimer: |
  # We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
#   
#   `\appendixfigures` needs to be added in front of appendix figures
#   `\appendixtables` needs to be added in front of appendix tables
#   
#   Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
editor_options: 
  markdown: 
    wrap: 72
---

@@

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2022]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Pecuchet2022 or @Kortsch2021?]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate the interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch 2021; @Gauzens2018] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.[WHAT ARE THE IMPLICATION OF THIS; NOT INCLUSDING BIOMASS; WHAT IS THE ASSUMPTION WE ARE MAKING; MAYBE WE NEED TO ADD A FEW WORDS ABOUT THIS IN THE DISCUSSION]

@@

Furthermore, we took a species’ habitat into account, which describes the physical position of a species within the ecosystem. Species were categorised as: 1) benthic, if a species lives on  the seafloor; 2) pelagic, if a species lives close to the surface; 3) benthopelagic, if it moves between and connects the mentioned environments; 4) demersal, if it lives and feeds on or near the bottom of the sea; and 5) land-based, if the consumer is not strictly aquatic but feeds predominantly on marine species. Species habitat affiliation was retrieved from @Jacob2011.

With the aim of studying the relationship between the interaction strength of the species (weighted property) and its unweighted properties we performed linear regression analyses between the log mean interaction strength and each of the mentioned unweighted properties. Thus we considered the interaction strength as the dependent variable and the given unweighted property as the independent variable, and obtained the coefficients (slope and intercept) for the linear model. Models were fitted using the least squares approach. We also explored the mean interaction strength distribution with the species habitat.

@@

To analyse the impact of species on food web stability, we performed extinction simulations deleting one species at a time, that is for every extinction, network size was reduced by one species only. After each extinction, we calculated the stability of the network minus the targeted species (489 nodes) and compared it with that of the whole network (490 nodes in total). To calculate stability, we used the mean of the real part of the maximum eigenvalue of the Jacobian matrix using randomized Jacobians, keeping the predator-prey sign structure fixed [@Allesina2008; @Grilli2016]. This stability index indicates a more stable food web when it is negative. We performed 1000 simulations for each species removal and obtained a mean maximum eigenvalue for each case. Finally, we statistically analysed this difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If the difference between the networks without and with the removed species is positive, then the stability of the food web is higher without the targeted species, and vice versa. A detailed description on the stability calculations can be found in the supplementary material.

To identify the species with a relatively high effect on food web stability, we plotted the results for each species extinction on food web stability, considering weighted (interaction strength) and unweighted properties (WHICH?), and species habitat.

@@

Our analyses show that the distribution of species interaction strength at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of interaction strength in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of interaction strength estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018]. <!--This finding reinforces the call for the inclusion of IS in food web studies to understand better the ecosystem functioning, and species and whole network responses to environmental perturbations.-->

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea in particular and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we have integrated information about weighted (interaction strength) and unweighted species properties, enabling a more complete assessment of the species’ role in the food web structure and function.

>>>>

---
title: New insights into the Weddell Sea ecosystem applying a quantitative network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2

correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web, which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on the species and trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted food web properties, and analysed species’ roles with respect to the stability of the food web. We found that the distribution of the interaction strength at the food web level is asymmetric, with many weak interactions and few strong ones. We detected a positive relationship between species mean interaction strength and two unweighted properties (i.e., trophic level and the total number of interactions). We also found that only a few species possess key positions in terms of food web stability. These species are characterised by high mean interaction strength, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In this study, we integrated unweighted and weighted food web information, enabling a more complete assessment of the ecosystem structure and function of the Weddell Sea food web. Our results provide new insights, which are important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

bibliography: ../WeddellSea.bib
running:
  title: New insights into the Weddell Sea food web
  author: Marina et al.
competinginterests: |
  The authors declare no competing interests.

availability:

codedata: |
    R code and data set used in this work is available at https://github.com/EcoComplex/WeddellSea

authorcontribution: |
  TIM and LAS: Conceptualization (lead); Data curation (lead); Formal analysis (lead); Methodology (lead); Coding (lead); Writing – original draft (lead); Writing – review and editing (lead). SK: Conceptualization (lead); Formal analysis (supporting); Methodology (supporting); Coding (supporting); Writing – original draft (supporting); Writing – review and editing (supporting).

acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}

output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article
editor_options: 
  markdown: 
    wrap: 72
---

@@

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2022]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Kortsch2021]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch 2021; @Gauzens2018] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.[WHAT ARE THE IMPLICATION OF THIS; NOT INCLUSDING BIOMASS; WHAT IS THE ASSUMPTION WE ARE MAKING; MAYBE WE NEED TO ADD A FEW WORDS ABOUT THIS IN THE DISCUSSION]

@@

Furthermore, we took species habitat affiliation into account, which describes the physical position of a species within the ecosystem. Species were categorised as: 1) benthic, if a species lives on the seafloor; 2) pelagic, if a species lives close to the surface; 3) benthopelagic, if it moves between and connects the aforementioned environments; 4) demersal, if it lives and feeds on, or near, the bottom of the sea; and 5) land-based, if the consumer is not strictly aquatic but feeds predominantly on marine species. Species habitat affiliations were retrieved from @Jacob2011.

To study the relationship between species mean interaction strength (weighted property) and the unweighted species properties, we performed linear regression analyses between the log mean interaction strength and each of the aforementioned unweighted properties. Thus, we considered the interaction strength as the dependent variable and the given unweighted property as the independent variable, and obtained the coefficients (slope and intercept) for the linear model. Models were fitted using the least squares approach. We also explored the mean interaction strength distribution with the species habitat.

@@

To analyse the impact of species on food web stability, we performed extinction simulations deleting one species at a time, that is for every extinction, network size was reduced by one species only. After each extinction, we calculated the stability of the network minus the removed species (489 nodes) and compared it with that of the whole network (490 nodes in total). To calculate stability, we used the mean of the real part of the maximum eigenvalue of the Jacobian matrix using randomized Jacobians, keeping the predator-prey sign structure fixed [@Allesina2008; @Grilli2016]. This stability index indicates a more stable food web when it is negative. We performed 1000 simulations for each species removal and obtained a mean maximum eigenvalue for each case. Finally, we statistically analysed this difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If the difference between the networks without and with the removed species is positive, then the stability of the whole food web is higher without the targeted species, in other words this species makes the network less stable. If the difference is negative, then the stability of the whole food web is lower without the targeted species, i.e., this species ahs a stabilizing effect. A detailed description on the stability calculations can be found in the supplementary material.

To identify the species with the highest effect on food web stability, we plotted the results of each species' extinction and its effect on food web stability, considering weighted (interaction strength) and unweighted properties (WHICH?), and species habitat. [THIS IS NOT 100% CLEAR; I NEED TO SEE THE FIGURES TO IMPROVE HERE, DON'T SEE THEM IN THIS FILE]

@@

Our analyses show that the distribution of species interaction strength at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of interaction strength in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of interaction strength estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018].

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea in particular, and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we integrated information on weighted (interaction strength) and unweighted species properties, enabling a more complete assessment of the species’ role realted food web structure and stability, and allowing us to identify species and their characteristics which can have a destabilising or stabilising effect on the food web.
